### 2024-08-07 meetings/MeetingNotes_07082024.Rmd
meeting notes 7 August

---
title: "Meeting Metaviromics 7 August 2024"
author: "laura"
date: "2024-08-07"
output: github_document
---

## Meeting Notes

Attendees: Sophie Gryseels, Magdalini Bletsa, Philippe Selhorst, Jayna Raghwani, Christina Faust, Laura Bergner

Sophie - challenging museum specimens and human samples (FFPE)\
Wildlife trafficking, wild meat sampling\
TE as screening tool

Magda - metagenomics, hepaciviruses\
FFPE - HIV using hybrid capture\
Pathogens carried by migratory birds\
Ancient DNA lab, hybrid capture for Hep B in ancient human samples

Philippe - metagenomic protocols for Nanopore

Less good experience with Twist\
Using depletion instead of enrichment\
Mastomys, positive selection worked less well than depletion, but don't know about what was in the panel or what was done by the company\
SISPA generates short pieces so haven't been using adaptive sampling

Ordering probes family by family\
CoV, ParamyxoV, Flavis\
Tested on known positive samples\
Worked well with Illumina seq\
As you combine probes how does that affect efficiency

RNA viruses mainly of interest\
Balancing depth vs breadth\
38 virus families, most RNA\
Drod vir database & VIRION\
Include viruses from metagenomic surveys as well as some from databases\
460 virus genomes

They are doing ~400 genomes per family\
Flavis started with 8000, after clustering had 800 genomes used to design probes\
Similar for CoV and ParamyxoV

Not sure about combining DNA and RNA viruses\
Better to use nuclease treatment or sensitivity is lost\
Need to split workflow\
Could use DNA transcripts in a RNA workflow but won't get genomes

They are using myBaits, RNA probes\
50-60K probes per family\
Smaller tiling\
Have not combined them\
150K probes in the combined set, unsure how that will affect results

Less genomes more families vs more families less genomes?

Current panels cluster at 90% identity\
Can capture different strains of the same virus (in HIV)

Have only done known positive samples so far\
Real bio controls would be better than something synthetic\
Spike in controls for non rodent pathogen on panel\
They've done spike in control in just pure water\
We are wanting also to quantify variation between samples\
Importance of negative controls for monitoring cross sample contamination

They have also found more viruses with shrews\
They also found a CoV but having difficulty assembling full genome\
They were sequencing from organs but have gotten nearly full genomes from shrew faecal samples in Glasgow\
Combine data from CoVs across teams showing geographical spread

Timeline Glasgow: doing some more metagenomic sequencing\
Start to develop panel in ~December\
First run test in new year\
Test out on samples from Oxford\
Eventually do sequencing in country

Problem with targeting what is already known in rodents, might miss out on surprising/new\
Per family - mammalian all included\
Main question is around host switching so don't want to exclude anything

Could use ancestral reconstruction to look for viruses that don't exist in contemporary dataset, include those in bait design\
Doing wider phylo analysis and ancestral reconstruction to guess at diversity

To make it universal, extra RNA and do RT with overhang instead of random primer. Use template switching RT step then amplify from those overhung sites (similar to S9N)\
Then could design downstream steps including barcodes to be more platform agnostic\
Potential to combine S9N, hybrid barcodes and twist hyb?\
Could potentially be more sensitive

Crispr based negative selection to get rid of host reads\
Can use with template switching\
Can be used to get rid of host RNA\
Developed for Mastomys and humans\
What is the most abundant RNA\
Wanting to make a central repository for Crispr panels so this data can be shared\
Going to make an Aedes/Culex panel\
Could even combined with target enrichment\
Downside is that you need one probe or Cas9 module for each molecule\
Originally used in transcriptomics on cells to deplete rRNA\
More difficult in DNA samples\
Expensive for an initial order but includes a lot\
Order two primers, anneal and elongate\
One ds 200 bp molecule per target

Benchmarking - on the virus level similar but on the depletion level it works better\
Compared to commercial rRNA kits\
More sensitive compared to generic hosts included in commercial panels\
5 euros per sample\
Not yet published

Get away from commercial assays to a more flexible and modular approach\
Discovery funding - apply for something together in the future?