### 2025-01-22 manuscript_figures.Rmd
Updating supp figures

---
title: "target_offtarget"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

This is an [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code.

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*.

```{r}
##all viruses - target and off target

library(tidyverse)
library(ggpubr)
library(tidyplots)
library(ggthemes)
library(scales)

```

### Figure 1 - linear model

#### Read-in data
```{r}
#Linear model for TE data - separated by virus
#November 2024

#viral read depth
#use bowtie2 data

library(tidyverse)
library(broom)
library(boot)

####read counts/normalised read counts####

#metadata

metadata <-
  read.csv(
    "metadata/sampleIDs_TESpikeIn.csv",
    header = TRUE
  )

metadata2 <- metadata |>
  select(Sample.ID, Number.of.read.pairs..quality.adaptor.trimmed.) |>
  rename(Sample_id = Sample.ID) |>
  rename(QC_reads = Number.of.read.pairs..quality.adaptor.trimmed.)

#import and combine read count files
#deduplicated and non-deduplicated

counts_dedup_bt <-
  read.table(
    "data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
    sep = "\t",
    header = TRUE
  )

counts_nodedup_bt <-
  read.table(
    "data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
    sep = "\t",
    header = TRUE
  )

counts_bt_all <- rbind(counts_dedup_bt, counts_nodedup_bt)

counts_reads <- left_join(counts_bt_all, metadata2, by = "Sample_id")

####read depths####

#import and combine read depth files

depth_dedup_bt <-
  read.table(
    "data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
    sep = "\t",
    header = TRUE
  )

depth_nodedup_bt <-
  read.table(
    "data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
    sep = "\t",
    header = TRUE
  )

```

#### Format data
```{r}

depths_bt_all <- rbind(depth_dedup_bt, depth_nodedup_bt)

depths_reads <- left_join(depths_bt_all, metadata2, by = "Sample_id") |>
  mutate(genome_structure = case_when((
    virus == "Human_adenovirus_40" |
      virus == "Human_betaherpesvirus"
  ) ~ "DNA",
  ,
  .default = "RNA"
  ))

####linear model for spike in viruses - mean read depth####

depths_reads_sub <- depths_reads |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  filter(type == "dedup_TE") |>
  mutate(log_depth = log10(mean_depth)) |>
  mutate(log_viral_load = log10(Viral.load))

#inspect data
hist(depths_reads_sub$mean_depth)
hist(depths_reads_sub$Viral.load)
hist(depths_reads_sub$log_depth)
hist(depths_reads_sub$log_viral_load)

summary(depths_reads_sub$log_depth)
summary(depths_reads_sub$log_viral_load)

```

#### model
```{r}
#all viruses together

lm1 <-
  glm(log_depth ~ log_viral_load * virus,
      data = depths_reads_sub,
      family = "gaussian")

summary(lm1)

glm.diag.plots(lm1)

pred <- predict(lm1, type = "response")

rsq <- function (x, y) {
  cor(x, y) ^ 2
}

rsq(depths_reads_sub$log_depth, pred)

##split by viruses

cols <- c("#4477AA",
          "#66CCEE",
          "#228833",
          "#CCBB44",
          "#EE6677",
          "#AA3377")

facet_names <- c(
  "Human_adenovirus_40" = "Human adenovirus 40",
  "Human_betaherpesvirus" = "Human betaherpesvirus",
  "Human_respiratory_syncytial_virus" = "Human respiratory syncytial virus",
  "Influenza_B_virus" = "Influenza B virus",
  "Mammalian_orthoreovirus3" = "Mammalian orthoreovirus 3",
  "Zika_virus" = "Zika virus"
)

virus <- (unique(depths_reads_sub$virus)) %>%
  rep(., each = 2) %>%
  data.frame()

list_models <- 
  depths_reads_sub |>
  group_split(virus) |>
  map( ~ lm(log_depth ~ log_viral_load, data = .))

lm_tidy <- 
  map(list_models, broom::tidy) %>%
  do.call(rbind.data.frame, .) %>%
  cbind(virus, .) %>%
  rename(virus = ".") %>%
  select(virus, term, estimate) %>%
  pivot_wider(names_from = "term", values_from = "estimate")

lm_summary <- depths_reads_sub |>
  group_by(virus) |>
  summarise(
    Intercept = lm(log_depth ~ log_viral_load)$coefficients[1],
    Coeff_x1 = lm(log_depth ~ log_viral_load)$coefficients[2],
    R2 = summary(lm(log_depth ~ log_viral_load))$r.squared,
    pvalue = summary(lm(log_depth ~ log_viral_load))$coefficients["log_viral_load", 4]
  )

lm_combined <- 
  left_join(lm_tidy, lm_summary, by = "virus")

```

#### Make plot
```{r, message=FALSE, warning=FALSE, echo=FALSE}
ggplot() +
  geom_point(data = depths_reads_sub,
             mapping = aes(Viral.load, mean_depth, color = virus)) +
  geom_abline(data = lm_combined, aes(intercept = `(Intercept)`, slope = log_viral_load), linetype="dotted") +
  geom_text(data = lm_combined,
            colour = "black",
            aes(
              x = 6*10000,
              y = 2,
              label = round(Coeff_x1, digits = 2)),
            size = 3.5) +
  geom_text(data = lm_combined,
            colour = "black",
            aes(x = 1.5*10000, 
                y = 2, 
                label = "Slope ="),
            size = 3.5) +
  geom_text(data = lm_combined,
            colour = "black",
            aes(
              x = 6*10000,
              y = 6,
              label = round(R2, digits = 2)),
            size = 3.5) +
  geom_text(data = lm_combined,
            colour = "black",
            aes(x = 2*10000, 
                y = 6, 
                label = "R2 ="),
            size = 3.5) +
  facet_wrap( ~ virus, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray"),
    axis.title.y = element_text(size = 10),
    legend.title = element_blank()
  ) +
    xlab("Log (Viral load)") + 
  ylab("Log (Mean Read Depth)") +
  scale_color_manual(values = cols) +
  guides(color = FALSE) +
  scale_x_log10(labels = scales::trans_format("log10", scales::label_math())) + 
    scale_y_log10(labels = scales::trans_format("log10", scales::label_math()))

```

#### Final plot
```{r}
#ggsave("figures/compare_spike_ins_atcc/model_readdepth_dedup.png")

ggsave("figures/manuscript_figure_2025/PDF/Figure_1.pdf",
       width = 8,
       height = 5)

ggsave("figures/manuscript_figure_2025/PNG/Figure_1.png",
       width = 8,
       height = 5)

```

### --------------------

## >>>> now redundant? - Figure 1: plots of TE experiment data - read counts and read depths

### Read-in data
```{r}
#ATCC genomes updated version
#October 2024

#viral load vs read count normalised using different methods, comparing deduplicated and non-deduplicated
#use bowtie2 data

####read counts/normalised read counts####

#metadata

metadata <- read.csv("metadata/sampleIDs_TESpikeIn.csv", header = TRUE)

metadata2 <- metadata |>
  select(Sample.ID, Number.of.read.pairs..quality.adaptor.trimmed.) |>
  rename(Sample_id = Sample.ID) |>
  rename(QC_reads = Number.of.read.pairs..quality.adaptor.trimmed.)

#import and combine read count files
#deduplicated and non-deduplicated

counts_dedup_bt <-
  read.table(
    "data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
    sep = "\t",
    header = TRUE
  )

counts_nodedup_bt <-
  read.table(
    "data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
    sep = "\t",
    header = TRUE
  )

counts_bt_all <- rbind(counts_dedup_bt, counts_nodedup_bt)

counts_reads <- left_join(counts_bt_all, metadata2, by = "Sample_id")

#normalise by both raw read count and genome length - should be the same as mean read depth

counts_reads_norm <- counts_reads |>
  mutate(norm_counts1 = matched / QC_reads) |>
  mutate(norm_counts2 = matched / QC_reads / length) |>
  mutate(norm_counts3 = matched / length) |>
  mutate(genome_structure = case_when((
    virus == "Human_adenovirus_40" |
      virus == "Human_betaherpesvirus"
  ) ~ "DNA",
  .default = "RNA"
  ))

```

### Format data
```{r}
#change labels in facet plots

facet_names <- c("dedup_TE" = "Deduplicated",
                 "nodedup_TE" = "Non-Deduplicated")

cols <- c("#4477AA",
          "#66CCEE",
          "#228833",
          "#CCBB44",
          "#EE6677",
          "#AA3377")
```

### Make plots
```{r}

#plot viral read counts (non normalised)

read_count <- 
  counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log10(matched),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  ylab("log10(Viral Reads)") +
  xlab("Log (Viral load)") + 
  scale_x_log10(labels = scales::trans_format("log10", scales::label_math()))

#ggsave("figures/compare_spike_ins_atcc/read_counts.pdf",width=8,height=6)

#normalise by raw read count

read_count_norm1 <- counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log10(norm_counts1),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  ylab("log10(viral reads/cleaned reads)") +
  xlab("Log (Viral load)") + 
  scale_x_log10(labels = scales::trans_format("log10", scales::label_math()))

#normalise by raw read count and genome length

read_count_norm2 <- 
  counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = (Viral.load),
    y = log10(norm_counts2),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  ylab("Log (viral reads/cleaned reads/genome length)") +
  xlab("Log (Viral load)") + 
  scale_x_log10(labels = scales::trans_format("log10", scales::label_math()))

```

### plot just the two lefthand panels from the original figure
```{r, warning=FALSE, echo=FALSE, message=FALSE, fig.height=10, fig.width=8}

ggarrange(read_count,
          read_count_norm2,
          nrow = 2,
          common.legend = TRUE)

#ggsave("figures/compare_spike_ins_atcc/log_read_count_depth_2panels.png",width=10,height=7)

```

### plot viral read counts (non normalised)
```{r}

##deduplicated only

read_count_dedup <- 
  counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  filter(type == "dedup_TE") |>
  ggplot(aes(
    x = Viral.load,
    y = log10(matched),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE, method = "lm") +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  ylab("Log (Viral Reads)") +
  xlab("Log (Viral load)") + 
  scale_x_log10(labels = scales::trans_format("log10", scales::label_math()))

#normalise by raw read count and genome length

read_count_norm2_dedup <- 
  counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  filter(type == "dedup_TE") |>
  ggplot(aes(
    x = Viral.load,
    y = log10(norm_counts2),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE, method = "lm") +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  ylab("Log (viral reads/cleaned reads/genome length)") +
  xlab("Log (Viral load)") + 
  scale_x_log10(labels = scales::trans_format("log10", scales::label_math()))
```

### ***Final plot***
```{r, warning=FALSE, message=FALSE, echo=FALSE, fig.width=8, fig.height=10}
#plot just the two lefthand panels from the original figure

ggarrange(
  read_count_dedup,
  read_count_norm2_dedup,
  nrow = 2,
  common.legend = TRUE,
  align = "hv"
)

# ggsave("figures/compare_spike_ins_atcc/log_read_count_depth_2panels_dedup.png",
#        width=10,
#        height=7)

ggsave("figures/manuscript_figure_2025/PDF/Figure_S1.pdf",
       width=7,
       height=8)

ggsave("figures/manuscript_figure_2025/PNG/Figure_S1.png",
       width=7,
       height=8)

```

### --------------------

## Figure S1

### Read depths
```{r}
####read depths####

#import and combine read depth files

depth_dedup_bt <-
  read_tsv(
    "data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv"
  )

depth_nodedup_bt <-
  read_tsv(
    "data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv"
  )

depths_bt_all <- rbind(depth_dedup_bt, depth_nodedup_bt)

depths_reads <- 
  depths_bt_all |>
  left_join(metadata2, by = "Sample_id") |>
  mutate(genome_structure = case_when((
    virus == "Human_adenovirus_40" |
      virus == "Human_betaherpesvirus"
  ) ~ "DNA",
  .default = "RNA"
  )) |>
  rename(Viral.load = `Viral load`)

```

### Make plots
```{r}

read_depths <- 
  depths_reads |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log10(mean_depth),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  ylab("Log (mean read depth)") +
  xlab("Log (Viral load)") + 
  scale_x_log10(labels = scales::trans_format("log10", scales::label_math()))

#ggsave("figures/compare_spike_ins_atcc/mean_depth.pdf",width=8,height=6)

```

### ***Final plot***
```{r, warning = FALSE, echo = FALSE, message = FALSE}

ggarrange(
  read_count,
  read_count_norm1,
  read_count_norm2,
  read_depths,
  nrow = 2,
  ncol = 2,
  common.legend = TRUE,
  align = "hv"
)

#ggsave("figures/compare_spike_ins_atcc/log_read_count_depth.png",width=10,height=7)

#ggsave("figures/manuscript_figures_pdf/FigureS1.pdf",width=10,height=7)

#apparently normalised read count using method 2 should not actually be the same

ggsave("figures/manuscript_figure_2025/PDF/Figure_S1.pdf",
       width = 12,
       height = 8)

ggsave("figures/manuscript_figure_2025/PNG/Figure_S1.png",
       width = 12,
       height = 8)

```

### --------------------

## Figure S2

### Read-in data / Make plots
```{r}
#metadata

metadata <-
  read.csv(
    "metadata/sampleIDs_TESpikeIn.csv",
    header = TRUE
  )

metadata2 <- metadata |>
  select(Sample.ID, Number.of.read.pairs..quality.adaptor.trimmed.) |>
  rename(Sample_id = Sample.ID) |>
  rename(QC_reads = Number.of.read.pairs..quality.adaptor.trimmed.)

#import and combine read count files
#deduplicated and non-deduplicated

counts_dedup_bt <-
  read.table(
    "data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
    sep = "\t",
    header = TRUE
  )

counts_nodedup_bt <-
  read.table(
    "data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
    sep = "\t",
    header = TRUE
  )

counts_bt_all <- rbind(counts_dedup_bt, counts_nodedup_bt)

counts_reads <- left_join(counts_bt_all, metadata2, by = "Sample_id")

#normalise by both raw read count and genome length - should be the same as mean read depth

counts_reads_norm <- counts_reads |>
  mutate(norm_counts1 = matched / QC_reads) |>
  mutate(norm_counts2 = matched / QC_reads / length) |>
  mutate(norm_counts3 = matched / length) |>
  mutate(genome_structure = case_when((
    virus == "Human_adenovirus_40" |
      virus == "Human_betaherpesvirus"
  ) ~ "DNA",
  .default = "RNA"
  ))

####compare library capture pool####

cols3 <- c("#228833", "#AA3377")

facet_names <- c("dedup_TE" = "Deduplicated",
                 "nodedup_TE" = "Non-Deduplicated")

metadata3 <- 
  metadata |>
  select(Sample.ID, Pool.for.sequencing) |>
  rename(Sample_id = Sample.ID) |>
  rename(Pool = Pool.for.sequencing)

counts_pool <- left_join(counts_reads_norm, metadata3, by = "Sample_id")

```

### Make plots
```{r}

pool_counts_sum <- 
  counts_pool |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = as.character(Viral.load),
    y = log10(matched),
    colour = Pool
  )) +
  geom_boxplot() +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_blank(),
    # axis.text.x = element_blank(),
    legend.title = element_blank(),
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  ylab("Log (Viral Reads)") + xlab("Log (Viral load)") +
  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2")) + 
  scale_x_discrete(labels = c(bquote(10^{2}), bquote(10^{3}), bquote(10^{5})))

# pool_readcounts_norm <- 
#   counts_pool |>
#   filter(Background != "p6") |>
#   filter(Background != "control") |>
#   ggplot(aes(x = virus, y = log10(norm_counts1))) +
#   geom_boxplot() +
#   facet_grid(Pool ~ type) +
#   theme_few() +
#   theme(
#     axis.title.x = element_blank(),
#     # axis.text.x = element_blank(),
#     axis.title.y = element_text(size = 10),
#     panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
#     panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
#   ) +
#   ylab("log10(viral reads/cleaned reads)")

pool_norm1_sum <- 
  counts_pool |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = as.character(Viral.load),
    y = log10(matched),
    colour = Pool
  )) +
  geom_boxplot() +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_blank(),
    # axis.text.x = element_blank(),
    legend.title = element_blank(),
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray") 
  ) +
  ylab("log10(viral reads/cleaned reads)") + xlab("Log (Viral load)") +
  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2")) + 
    scale_x_discrete(labels = c(bquote(10^{2}), bquote(10^{3}), bquote(10^{5})))

# pool_readcounts_norm2 <- 
#   counts_pool |>
#   filter(Background != "p6") |>
#   filter(Background != "control") |>
#   ggplot(aes(x = virus, y = log10(norm_counts2))) +
#   geom_boxplot() +
#   facet_grid(Pool ~ type) +
#   theme_few() +
#   theme(
#     axis.title.x = element_blank(),
#     # axis.text.x = element_blank(),
#     axis.title.y = element_text(size = 10),
#     panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
#     panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
#   ) +
#   ylab("log10(viral reads/cleaned reads/genome length)")

pool_norm2_sum <- 
  counts_pool |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = as.character(Viral.load),
    y = log10(matched),
    colour = Pool
  )) +
  geom_boxplot() +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_blank(),
    # axis.text.x = element_blank(),
    legend.title = element_blank(),
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  ylab("log10(viral reads/cleaned reads/genome length)") + xlab("Log (Viral load)") +
  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2")) + 
    scale_x_discrete(labels = c(bquote(10^{2}), bquote(10^{3}), bquote(10^{5})))

#import and combine read depth files

depth_dedup_bt <-
  read.table(
    "data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
    sep = "\t",
    header = TRUE
  )

depth_nodedup_bt <-
  read.table(
    "data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
    sep = "\t",
    header = TRUE
  )

depths_bt_all <- rbind(depth_dedup_bt, depth_nodedup_bt)

depths_reads <- 
  left_join(depths_bt_all, metadata2, by = "Sample_id") |>
  mutate(genome_structure = case_when((
    virus == "Human_adenovirus_40" |
      virus == "Human_betaherpesvirus"
  ) ~ "DNA",
  .default = "RNA"
  ))

depths_pool <- left_join(depths_reads, metadata3, by = "Sample_id")

# pool_readdepths <- 
#   depths_pool |>
#   filter(Background != "p6") |>
#   filter(Background != "control") |>
#   ggplot(aes(x = virus, y = log10(mean_depth))) +
#   geom_boxplot() +
#   facet_grid(Pool ~ type) +
#   theme_few() +
#   theme(
#     axis.title.x = element_blank(),
#     axis.text.x = element_text(angle = 45, hjust = 1),
#     axis.title.y = element_text(size = 10),
#     panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
#     panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
#   ) +
#   ylab("Log (mean read depth)")

pool_depths_sum <-
  depths_pool |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = as.character(Viral.load),
    y = log10(mean_depth),
    colour = Pool
  )) +
  geom_boxplot() +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_blank(),
    # axis.text.x = element_blank(),
    legend.title = element_blank(),
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  ylab("Log (mean read depth)") +
  xlab("Log (Viral load)") + 
  scale_x_discrete(labels = c(bquote(10^{2}), bquote(10^{3}), bquote(10^{5}))) +
  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2"))

```

### ***Final plot***
```{r}
ggarrange(
  pool_counts_sum,
  pool_norm1_sum,
  pool_norm2_sum,
  pool_depths_sum,
  nrow = 2,
  ncol = 2,
  common.legend = TRUE, 
  align = "hv"
)

#ggsave("figures/compare_spike_ins_atcc/pool_compare_viruses_sum.png",width=10,height=7)

ggsave("figures/manuscript_figure_2025/PDF/Figure_S2.pdf",
       width = 10,
       height = 7)

ggsave("figures/manuscript_figure_2025/PNG/Figure_S2.png",
       width = 10,
       height = 7)

```

### --------------------

## Figure S3
```{r}

##plots of read counts and viral read counts
#November 2024

#metadata

metadata <-
  read.csv(
    "metadata/sampleIDs_TESpikeIn.csv",
    header = TRUE
  )

metadata2 <- metadata |>
  select(Sample.ID, Number.of.read.pairs..quality.adaptor.trimmed.) |>
  rename(Sample_id = Sample.ID) |>
  rename(QC_reads = Number.of.read.pairs..quality.adaptor.trimmed.)

#import read count files

counts_dedup_bt <-
  read.table(
    "data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
    sep = "\t",
    header = TRUE
  )

counts_nodedup_bt <-
  read.table(
    "data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
    sep = "\t",
    header = TRUE
  )

#import viral reads mapped (to calculate proportions)

viral_reads_dedup <-
  read.csv(
    "data_TE/total_virus_mapped_reads_per_sample_dedup_atcc_ref_20241108.csv",
    header = TRUE
  )

viral_reads_nodedup <-
  read.csv(
    "data_TE/total_virus_mapped_reads_per_sample_nodedup_atcc_ref_20241108.csv",
    header = TRUE
  )

cols2 <- c("#BB5566", "#004488")

facet_names <- c("dedup_TE" = "Deduplicated",
                 "nodedup_TE" = "Non-Deduplicated")

reads_metadata_dedup <-
  left_join(counts_dedup_bt, metadata2, by = "Sample_id")

reads_viral_dedup <-
  left_join(reads_metadata_dedup, viral_reads_dedup, by = "Sample_id")

reads_metadata_nodedup <-
  left_join(counts_nodedup_bt, metadata2, by = "Sample_id")

reads_viral_nodedup <-
  left_join(reads_metadata_nodedup, viral_reads_nodedup, by = "Sample_id")

reads_viral_all <- rbind(reads_viral_dedup, reads_viral_nodedup)

reads_plot <- 
  reads_viral_all |>
  group_by(Background, Sample_id, Viral.load, type) |>
  summarise(
    total_reads = (QC_reads * 2),
    viral_reads = total_virus_reads,
    ATCC_reads = sum(matched),
    prop_ATCC = ATCC_reads / total_reads,
    prop_viral = total_virus_reads / total_reads,
    diff = viral_reads - ATCC_reads
  ) |>
  unique()

```

### Make plots
```{r, echo=FALSE, warning=FALSE, message=FALSE}

total_reads <- 
  reads_plot |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = (total_reads),
    colour = Background
  )) +
  geom_point() +
  geom_smooth(aes(group = Background), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_y_log10(labels = scales::trans_format("log10", scales::label_math())) + 
    scale_x_log10(labels = scales::trans_format("log10", scales::label_math())) + xlab("Viral load") +
  # ggtitle("All") +
  ylab("Total number of reads") +
  scale_color_manual(values = cols2, labels = c("P1", "P2"))

viral_reads <- 
  reads_plot |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = (viral_reads),
    colour = Background
  )) +
  geom_point() +
  geom_smooth(aes(group = Background), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_y_log10(labels = scales::trans_format("log10", scales::label_math())) + 
    scale_x_log10(labels = scales::trans_format("log10", scales::label_math())) + xlab("Viral load") +
  # ggtitle("Viral reads") +
  ylab("Number of viral reads") +
  scale_color_manual(values = cols2, labels = c("P1", "P2"))

ATCC_reads <- 
  reads_plot |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = (ATCC_reads),
    colour = Background
  )) +
  geom_point() +
  geom_smooth(aes(group = Background), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_y_log10(labels = scales::trans_format("log10", scales::label_math())) + 
    scale_x_log10(labels = scales::trans_format("log10", scales::label_math())) + xlab("Viral load") +
  ggtitle("Spike-in Viral Reads") +
  ylab("Spike-in Viral Reads") +
  scale_color_manual(values = cols2, labels = c("P1", "P2"))

prop_ATCC <- reads_plot |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(x = Viral.load, y = prop_ATCC, colour = Background)) +
  geom_point() +
  geom_smooth(aes(group = Background), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
    scale_x_log10(labels = scales::trans_format("log10", scales::label_math())) + xlab("Viral load") +
  ggtitle("Spike-in viral reads") +
  ylab("Proportion") +
  scale_color_manual(values = cols2, labels = c("P1", "P2"))

prop_viral <- reads_plot |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(x = Viral.load, y = prop_viral, colour = Background)) +
  geom_point() +
  geom_smooth(aes(group = Background), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
 scale_x_log10(labels = scales::trans_format("log10", scales::label_math())) + xlab("Viral load") +
  # ggtitle("Viral reads") +
  ylab("Proportion of viral reads") +
  scale_color_manual(values = cols2, labels = c("P1", "P2")) + scale_y_continuous(limits=c(0,1.0))

#ggarrange(total_reads,viral_reads,ATCC_reads,prop_viral,prop_ATCC,nrow=2,ncol=3,common.legend = TRUE)

```

### ***Final plot***
```{r, warning=FALSE, echo=FALSE, message=FALSE}

ggarrange(total_reads,
          viral_reads,
          prop_viral,
          nrow = 3,
          common.legend = TRUE,
          align = "hv")

#ggsave("figures/compare_spike_ins_atcc/backgrounds_reads.png")

ggsave("figures/manuscript_figure_2025/PDF/Figure_S3.pdf", 
       width = 8,
       height = 8)

ggsave("figures/manuscript_figure_2025/PNG/Figure_S3.png", 
       width = 8,
       height = 8)

####compare proportion viral reads per pool with shotgun####

# metadata3 <- metadata |>
#   select(Sample.ID,
#          Number.of.read.pairs..quality.adaptor.trimmed.,
#          Pool.for.sequencing) |>
#   rename(Sample_id = Sample.ID) |>
#   rename(QC_reads = Number.of.read.pairs..quality.adaptor.trimmed.)
# 
# reads_viral_dedup3 <-
#   full_join(metadata3, viral_reads_dedup, by = "Sample_id")
# 
# reads_viral_nodedup3 <-
#   full_join(metadata3, viral_reads_nodedup, by = "Sample_id")
# 
# reads_viral_dedup3 |>
#   group_by(Pool.for.sequencing) |>
#   summarise(
#     viral_reads = sum(total_virus_reads),
#     total_reads = sum(QC_reads * 2),
#     prop_viral = viral_reads / total_reads
#   )
# 
# polyomics <-
#   read.csv(
#     "data_polyomics/total_virus_mapped_reads_per_sample_dedup.csv"
#   )
# 
# #read data from polyomics_indexes_stefano document
# total_reads <- data.frame(total_reads = c(5942760, 7714275))
# 
# keeps <- c("RNA-Msp-p2", "RNA-Msp-p8")
# 
# polyomics_read_prop <- polyomics |>
#   filter(Sample_id %in% keeps)
# 
# polyomics_read_prop2 <- cbind(polyomics_read_prop, total_reads)
# 
# polyomics_read_prop2 |>
#   summarise(prop_viral = total_virus_reads / total_reads)

```

### --------------------

## No longer required? Figure S4

### Make plots
```{r}
####read counts/depths split by background####

#change labels in facet plots

facet_names_bg <- c(
  "dedup_TE" = "Deduplicated",
  "nodedup_TE" = "Non-Deduplicated",
  "p2" = "P1",
  "p8" = "P2"
)

#break down by Background x virus

background_counts_reads <- 
  counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log10(matched),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid(Background ~ type, labeller = as_labeller(facet_names_bg)) +
  theme_few() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_color_manual(
    values = cols,
    labels =
      c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
      )
  ) +
  scale_x_log10(labels = scales::label_log10()) +
  ylab("log10(Viral Reads)")

backgrounds_counts_reads_norm <- 
  counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log10(norm_counts1),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid(Background ~ type, labeller = as_labeller(facet_names_bg)) +
  theme_bw() +
  theme_few() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  scale_x_log10(label = scales::label_log10()) +
  ylab("log10(viral reads/cleaned reads)")

backgrounds_counts_reads_norm2 <- 
  counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log10(norm_counts2),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid(Background ~ type, labeller = as_labeller(facet_names_bg)) +
  theme_few() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 10),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  scale_x_log10(labels = scales::label_log10()) +
  ylab("log10(viral reads/cleaned reads/genome length)")

background_read_depths <- 
  depths_reads |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log10(mean_depth),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid(Background ~ type, labeller = as_labeller(facet_names_bg)) +
  theme_few() +
  theme(
    axis.title.x = element_blank(),
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray"),
    axis.title.y = element_text(size = 10),
    legend.title = element_blank()
  ) +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  scale_x_log10(label = scales::label_log10()) +
  ylab("log10(mean read depth)")

```

### ***Final plot*** 
```{r, warning=FALSE, message=FALSE, echo=FALSE, fig.height=6, fig.height=8}

ggarrange(
  background_counts_reads,
  backgrounds_counts_reads_norm,
  backgrounds_counts_reads_norm2,
  background_read_depths,
  nrow = 2,
  ncol = 2,
  common.legend = TRUE,
  align = "hv"
)

#ggsave("figures/compare_spike_ins_atcc/backgrounds_read_count_depth.png",width=10,height=7)

#ggsave("figures/manuscript_figures_pdf/FigureS4.pdf",width=10,height=7)

```

### --------------------
### --------------------

## >>> Figure 2: plots of TE experiment data - genome coverage & per site coverage

### Read-in data
```{r}

#plots of TE experiment data - genome coverage & per site coverage
#also separated by viruses
#ATCC genomes updated version
#October 2024

#use bowtie2 data

####genome coverage####

unzip(zipfile = "data_TE/TE_sequencing_experiment_readdepth_per_site_sample_and_virus_bowtie2_dedup_atcc_ref.tsv.zip", exdir = "data_TE/")

unzip(zipfile = "data_TE/TE_sequencing_experiment_readdepth_per_site_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv.zip", exdir = "data_TE/")

dedup_per_site <-
  read_tsv(
    "data_TE/TE_sequencing_experiment_readdepth_per_site_sample_and_virus_bowtie2_dedup_atcc_ref.tsv"
  )

nodedup_per_site <-
  read_tsv(
    "data_TE/TE_sequencing_experiment_readdepth_per_site_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv"
  )

file.remove("data_TE/TE_sequencing_experiment_readdepth_per_site_sample_and_virus_bowtie2_dedup_atcc_ref.tsv")

file.remove( "data_TE/TE_sequencing_experiment_readdepth_per_site_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv")

#add length column from read depth file to per site file

#metadata

metadata <-
  read.csv(
    "metadata/sampleIDs_TESpikeIn.csv",
    header = TRUE
  )

metadata2 <- 
  metadata |>
  select(Sample.ID, Number.of.read.pairs..quality.adaptor.trimmed.) |>
  rename(Sample_id = Sample.ID) |>
  rename(QC_reads = Number.of.read.pairs..quality.adaptor.trimmed.)

#import and combine read count files

#deduplicated and non-deduplicated

counts_dedup_bt <-
  read_tsv(
    "data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv"
  )

counts_nodedup_bt <-
  read_tsv(
    "data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv"
  )

counts_bt_all <- rbind(counts_dedup_bt, counts_nodedup_bt)

counts_reads <- left_join(counts_bt_all, metadata2, by = "Sample_id")

```

### Format data
```{r}
#normalise by both raw read count and genome length - should be the same as mean read depth

counts_reads_norm <- 
  counts_reads |>
  mutate(norm_counts1 = matched / QC_reads) |>
  mutate(norm_counts2 = matched / QC_reads / length) |>
  mutate(norm_counts3 = matched / length) |>
  mutate(genome_structure = case_when((
    virus == "Human_adenovirus_40" |
      virus == "Human_betaherpesvirus"
  ) ~ "DNA",
  .default = "RNA"
  ))

lengths <- 
  counts_reads_norm |>
  select(virus, length) |>
  distinct()

#calculate genome coverage

persite_coverage_dedup <- 
  dedup_per_site |>
  rename(Viral.load = `Viral load`) |>
  group_by(Sample_id, virus, Viral.load, Background, type) |>
  filter(coverage > 0) |>
  summarise(genome_coverage = n()) |>
  left_join(lengths) |>
  mutate(percent_coverage = genome_coverage / length)

persite_coverage_nodedup <- 
  nodedup_per_site |>
  rename(Viral.load = `Viral load`) |>
  group_by(Sample_id, virus, Viral.load, Background, type) |>
  filter(coverage > 0) |>
  summarise(genome_coverage = n()) |>
  left_join(lengths) |>
  mutate(percent_coverage = genome_coverage / length)

persite_coverage_both <-
  rbind(persite_coverage_dedup, 
        persite_coverage_nodedup)

coverage_labels <- c("0" = "Control",
                     "100" = "10^{2}",
                     "1000" = "10^{3}",
                     "10000" = "10^{5}",
                     "dedup_TE" = "Deduplicated",
                     "nodedup_TE"= "Non-Deduplicated")

coverage_labels2 <- c("0" = "Control",
                      "100" = "10^{2}",
                      "1000" = "10^{3}",
                      "10000" = "10^{5}")
```

### Make plots
```{r, echo = FALSE, warning=FALSE, message=FALSE, fig.height=4, fig.width=6}

cols2 <- c("#BB5566",
           "#004488")

cols4 <- c("#DDAA33",
           "#BB5566",
           "#004488")

persite_coverage_both$Viral.load <- factor(persite_coverage_both$Viral.load, labels = c("0",
    "10^{2}", 
    "10^{3}", 
    "10^{5}"
    ))

##are the deduplicated and non deduplicated datasets identical??
##if they are the same can maybe just show one
#show deduplicated

persite_coverage_both |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  filter(type == "dedup_TE") |>
  ggplot(aes(x = virus, y = percent_coverage, colour = Background)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_dodge(width = .75)) +
  facet_grid(~as.character(Viral.load), labeller = label_parsed) +
  theme_few() +
  ylab("Genome coverage") +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 30, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_color_manual(values = cols2, labels = c("P1", "P2")) +
  scale_x_discrete(
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  scale_y_continuous(limits = c(0, 1))

```

### ***Final plot***
```{r, echo = FALSE, warning=FALSE, message=FALSE}
#ggsave("figures/compare_spike_ins_atcc/genome_cov_deduponly.png",width=15,height=6)

ggsave(
  "figures/manuscript_figure_2025/PDF/Figure_2.pdf",
  width = 15,
  height = 6
)

ggsave(
  "figures/manuscript_figure_2025/PNG/Figure_2.png",
  width = 15,
  height = 6
)

```

### --------------------

### --------------------

### Figure S4
```{r}

persite_both <- rbind(dedup_per_site, nodedup_per_site)

persite_reads <- left_join(persite_both, metadata2, by = "Sample_id")

persite_norm <- persite_reads |>
  mutate(norm_cov = coverage / QC_reads)

coverage_labels3 <-c("100" = "1e+02",
                    "1000" = "1e+03",
                    "100000" = "1e+05",
                    "p2" = "P1",
                    "p8"= "P2")

persite_norm$`Viral load` <- 
  factor(persite_norm$`Viral load`, 
         labels = c("0",
                    "10^{2}", 
                    "10^{3}",
                    "10^{5}"))

persite_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  filter(type == "dedup_TE") |>
  filter(virus == "Human_respiratory_syncytial_virus") |>
  rename(Viral.load = `Viral load`) |>
  ggplot(aes(x = site, y = coverage, fill = Background)) +
  geom_col() +
  facet_wrap(Viral.load ~ Sample_id, label = label_parsed) +
  ylab("Coverage") +
  ggtitle("Human RSV (deduplicated)") +
  guides(fill = guide_legend(title = "Background")) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 30, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_y_log10() +
  scale_fill_manual(values = cols2, labels = c("P1", "P2"))
```

#### ***Final plot***
```{r}
ggsave("figures/manuscript_figure_2025/PDF/Figure_S4.pdf",
       width=15,
       height=10)

ggsave("figures/manuscript_figure_2025/PNG/Figure_S4.png",
       width=15,
       height=10)
```

### --------------------

### Figure S5
```{r, warning=FALSE, message=FALSE, echo=FALSE}

persite_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  filter(type == "dedup_TE") |>
  filter(virus == "Zika_virus") |>
  rename(Viral.load = `Viral load`) |>
  ggplot(aes(x = site, y = coverage, fill = Background)) +
  geom_col() +
  facet_wrap(Viral.load ~ Sample_id, label = label_parsed) +
  ylab("Coverage") +
  ggtitle("Zika virus (deduplicated)") +
  guides(fill = guide_legend(title = "Background")) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 30, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_y_log10() +
  scale_fill_manual(values = cols2, labels = c("P1", "P2"))

```

####***Final plot***
```{r}

ggsave("figures/manuscript_figure_2025/PDF/Figure_S5.pdf",
       width=15,
       height=10)

ggsave("figures/manuscript_figure_2025/PNG/Figure_S5.png",
       width=15,
       height=10)

```

### --------------------

### Figure S6
```{r, warning=FALSE, message=FALSE, echo=FALSE}

persite_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  filter(type == "dedup_TE") |>
  filter(virus == "Human_adenovirus_40") |>
  rename(Viral.load = `Viral load`) |>
  ggplot(aes(x = site, y = coverage, fill = Background)) +
  geom_col() +
  facet_wrap(Viral.load ~ Sample_id, label = label_parsed) +
  ylab("Coverage") +
  ggtitle("Human Adenovirus (deduplicated)") +
  guides(fill = guide_legend(title = "Background")) +
 theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 30, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_y_log10() +
  scale_fill_manual(values = cols2, labels = c("P1", "P2"))

```

####***Final plot***
```{r}

ggsave("figures/manuscript_figure_2025/PDF/Figure_S6.pdf",
       width=15,
       height=10)

ggsave("figures/manuscript_figure_2025/PNG/Figure_S6.png",
       width=15,
       height=10)

```

### --------------------

### Figure S7
```{r}
HBV <- 
persite_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  filter(type == "dedup_TE") |>
  filter(virus == "Human_betaherpesvirus") |>
  rename(Viral.load = `Viral load`) |>
  ggplot(aes(x = site, y = coverage, fill = Background)) +
  geom_col() +
  facet_wrap(Viral.load ~ Sample_id, label = label_parsed) +
  ylab("Coverage") +
  ggtitle("Human Betaherpesvirus (deduplicated)") +
  guides(fill = guide_legend(title = "Background")) +
 theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 30, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_y_log10() +
  scale_fill_manual(values = cols2, labels = c("P1", "P2"))

```

####***Final plot***
```{r}

ggsave(plot = HBV, file="figures/manuscript_figure_2025/PDF/Figure_S7.pdf",
       width=15,
       height=10,
       dpi = 600)

ggsave(plot = HBV, file="figures/manuscript_figure_2025/PNG/Figure_S7.png",
       width=15,
       height=10)

```

### --------------------

### Figure S8
```{r}
#have to do these differently due to segmentation

FLU_ME1 <- persite_norm |>
  filter(virus == "Influenza_B_virus") |>
  filter(type == "dedup_TE") |>
  filter(Background == "p2") |>
  rename(Viral.load = `Viral load`) |>
  ggplot(aes(
    x = site,
    y = coverage,
    fill = as.character(Viral.load)
  )) +
  geom_col() +
  facet_grid(seg ~ Sample_id) +
  ylab("Coverage") +
  ggtitle("Influenza B virus (Background 1 deduplicated)") +
  guides(fill = guide_legend(title = "Viral load")) +
 theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 30, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_y_log10() +
  scale_fill_manual(values = cols4, label = c(expression("10"^"2"), expression("10"^"3"), expression("10"^"5")))

FLU_ME2 <- persite_norm |>
  filter(virus == "Influenza_B_virus") |>
  filter(type == "dedup_TE") |>
  filter(Background == "p2") |>
  rename(Viral.load = `Viral load`) |>
  ggplot(aes(
    x = site,
    y = coverage,
    fill = as.character(Viral.load)
  )) +
  geom_col() +
  facet_grid(seg ~ Sample_id) +
  ylab("Coverage") +
  ggtitle("Influenza B virus (Background 2 deduplicated)") +
  guides(fill = guide_legend(title = "Viral load")) +
 theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 30, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_y_log10() +
  scale_fill_manual(values = cols4, label = c(expression("10"^"2"), expression("10"^"3"), expression("10"^"5")))

ggarrange(FLU_ME1, FLU_ME2, ncol = 2, common.legend = TRUE)

```

####***Final plot***
```{r}

ggsave("figures/manuscript_figure_2025/PDF/Figure_S8.pdf",
       width=20,
       height=12,
       dpi=600)

ggsave("figures/manuscript_figure_2025/PNG/Figure_S8.png",
       width=20,
       height=12)
```

### --------------------

### Figure S9
```{r}

REO_ME1 <- 
  persite_norm |>
  filter(virus == "Mammalian_orthoreovirus3") |>
  filter(type == "dedup_TE") |>
  filter(Background == "p2") |>
  rename(Viral.load = `Viral load`) |>
  ggplot(aes(
    x = site,
    y = coverage,
    fill = as.character(Viral.load)
  )) +
  geom_col() +
  facet_grid(seg ~ Sample_id) +
  ylab("log10(Coverage)") +
  ggtitle("Mammalian orthoreovirus 3 (Background 1 deduplicated)") +
  guides(fill = guide_legend(title = "Viral load")) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 30, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top",
    panel.grid.major = element_line(linewidth = 0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth = 0.1, color = "lightgray")
  ) +
  scale_y_log10() +
  scale_fill_manual(values = cols4,
                    label = c(
                      expression("10" ^ "2"),
                      expression("10" ^ "3"),
                      expression("10" ^ "5")
                    ))

REO_ME2 <- 
  persite_norm |>
  filter(virus == "Mammalian_orthoreovirus3") |>
  filter(type == "dedup_TE") |>
  filter(Background == "p8") |>
  rename(Viral.load = `Viral load`) |>
  ggplot(aes(
    x = site,
    y = coverage,
    fill = as.character(Viral.load)
  )) +
  geom_col() +
  facet_grid(seg ~ Sample_id) +
  ylab("log10(Coverage)") +
  ggtitle("Mammalian orthoreovirus 3 (Background 2 deduplicated)") +
  guides(fill = guide_legend(title = "Viral load")) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 30, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top",
    panel.grid.major = element_line(linewidth = 0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth = 0.1, color = "lightgray")
  ) + scale_y_log10() +
  scale_fill_manual(values = cols4,
                    label = c(
                      expression("10" ^ "2"),
                      expression("10" ^ "3"),
                      expression("10" ^ "5")
                    ))

ggarrange(REO_ME1, REO_ME2, ncol = 2, common.legend = TRUE)

```

####***Final plot***
```{r}
ggsave("figures/manuscript_figure_2025/PDF/Figure_S9.pdf",
       width=20,
       height=12,
       dpi=600)

ggsave("figures/manuscript_figure_2025/PNG/Figure_S9.png",
       width=20,
       height=12,
       dpi=600)
```

### --------------------

## >>>> Figure 3: all viruses - target and off target

### Read-in data
```{r}

# read in metadata

metadata <-
  read.csv("metadata/sampleIDs_TESpikeIn.csv", header = TRUE)

metadata2 <-
  metadata |>
  select(
    Sample.ID,
    Background.sample,
    Viral.load,
    Number.of.read.pairs..quality.adaptor.trimmed.
  ) |>
  rename(Sample_id = Sample.ID) |>
  rename(QC_reads = Number.of.read.pairs..quality.adaptor.trimmed.)

#import viral reads mapped (to calculate proportions)

viral_reads_dedup <-
  read.csv(
    "data_TE/total_virus_mapped_reads_per_sample_dedup_atcc_ref_20241108.csv",
    header = TRUE
  )

viral_reads_nodedup <-
  read.csv(
    "data_TE/total_virus_mapped_reads_per_sample_nodedup_atcc_ref_20241108.csv",
    header = TRUE
  )

#import contingency tables

contingency_dedup <-
  read.csv(
    "data_TE/contingency_table_mapped_virus_reads_per_family_genus_sample_dedup_atcc_ref_20241106.csv",
    header = TRUE
  )

contingency_nodedup <-
  read.csv(
    "data_TE/contingency_table_mapped_virus_reads_per_family_genus_sample_nodedup_atcc_ref_20241106.csv",
    header = TRUE
  )

contaminants <-
  c("Betacoronavirus", "Alphainfluenzavirus", "Gammaretrovirus")

```

### Format data
```{r}

# pivot longer  
dedup_long <-
  contingency_dedup |>
  filter(!genus %in% contaminants) |>
  filter(genus != "NA") |>
  pivot_longer(cols = A:P,
               names_to = "Sample_id",
               values_to = "count") |>
  mutate(
    target = case_when(
      genus == "Cyclovirus" ~ "off_target",
      genus == "Cardiovirus" ~ "off_target",
      genus == "Kobuvirus" ~ "off_target",
      .default = "on_target"
    )
  ) |>
  mutate(genome_structure = case_when((genus == "Mastadenovirus" |
                                         genus == "Cytomegalovirus") ~ "DNA",
                                      .default = "RNA"
  ))

no_dedup_long <- contingency_nodedup |>
  filter(!genus %in% contaminants) |>
  filter(genus != "NA") |>
  pivot_longer(cols = A:P,
               names_to = "Sample_id",
               values_to = "count") |>
  mutate(
    target = case_when(
      genus == "Cyclovirus" ~ "off_target",
      genus == "Cardiovirus" ~ "off_target",
      genus == "Kobuvirus" ~ "off_target",
      .default = "on_target"
    )
  ) |>
  mutate(genome_structure = case_when((genus == "Mastadenovirus" |
                                         genus == "Cytomegalovirus") ~ "DNA",
                                      .default = "RNA"
  ))

metadata_dedup <- left_join(dedup_long, metadata2, by = "Sample_id")

dedup_viral_reads <-
  left_join(metadata_dedup, viral_reads_dedup, by = "Sample_id") |>
  mutate(total_reads = QC_reads * 2) |>
  mutate(prop_total = count / total_reads) |>
  mutate(prop_viral = count / total_virus_reads)

metadata_no_dedup <- left_join(no_dedup_long, metadata2, by = "Sample_id")

nodedup_viral_reads <-
  left_join(metadata_no_dedup, viral_reads_nodedup, by = "Sample_id") |>
  mutate(total_reads = QC_reads * 2) |>
  mutate(prop_total = count / total_reads) |>
  mutate(prop_viral = count / total_virus_reads)

```

### Make plots
```{r, warning=FALSE, message=FALSE, echo=FALSE}

facet_names <- c(
  "Msp_p2" = "P1",
  "Msp_p8" = "P2",
  "off_target" = "Off target viruses",
  "on_target" = "On target viruses"
)

cols <-
  c(
    "#CCBB44",
    "#332288",
    "#EE7733",
    "#66CCEE",
    "#882255",
    "#4477AA",
    "#AA3377",
    "#228833",
    "#EE6677"
  )

dedup_reads <- 
  dedup_viral_reads |>
  filter(Background.sample != "Neg_control") |>
  filter(Background.sample != "Msp_p6") |>
  ggplot(aes(
    x = Viral.load,
    y = log10(count),
    colour = genus
  )) +
  geom_point() +
  geom_smooth(aes(group = genus, linetype = genome_structure), se = FALSE) +
  facet_grid(target ~ Background.sample, labeller = as_labeller(facet_names)) +
  theme_few() + 
  scale_x_log10() +
  ylab("Log (Viral Reads)") + 
  xlab("Log (Viral load)") + 
  scale_color_manual(values = cols) +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 30, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  )

dedup_prop_viral <-
  dedup_viral_reads |>
  filter(Background.sample != "Neg_control") |>
  filter(Background.sample != "Msp_p6") |>
  ggplot(aes(x = Viral.load, y = prop_viral, colour = genus)) +
  geom_point() +
  geom_smooth(aes(group = genus, linetype = genome_structure), se = FALSE) +
  facet_grid(target ~ Background.sample, labeller = as_labeller(facet_names)) +
  theme_few() + 
  scale_x_log10() +
  scale_y_log10() +
  ylab("Proportion of viral reads") +
  xlab("Log (Viral load)") + 
  scale_color_manual(values = cols) +
    theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 30, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  )

dedup_prop_all <-
  dedup_viral_reads |>
  filter(Background.sample != "Neg_control") |>
  filter(Background.sample != "Msp_p6") |>
  ggplot(aes(x = Viral.load, y = prop_total, colour = genus)) +
  geom_point() +
  geom_smooth(aes(group = genus, linetype = genome_structure), se = FALSE) +
  facet_grid(target ~ Background.sample, labeller = as_labeller(facet_names)) +
  theme_few() + 
  scale_x_log10() +
  scale_y_log10(limits=c(5e-8,1)) +
  ylab("Proportion of total reads") +
  xlab("Log (Viral load)") + 
  scale_color_manual(values = cols) +
    theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 30, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  )

```

### Dedup -- viral reads, prop all reads, and prop viral reads
```{r, warning=FALSE, message=FALSE, echo=FALSE, fig.width=8, fig.height=12}

ggarrange(
  dedup_reads,
  dedup_prop_all,
  dedup_prop_viral,
  nrow = 3,
  common.legend = TRUE,
  align = "hv"
)

```

### ***Final plot***

```{r, warning=FALSE, message=FALSE, echo=FALSE, fig.width=7, fig.height=8}

#ggsave("figures/compare_spike_ins_atcc/target_offtarget_dedup.png",width=8,height=10)

##just plot viral reads and prop viral

ggarrange(
  dedup_reads,
  dedup_prop_viral,
  nrow = 2,
  common.legend = TRUE,
  align = "hv"
)

ggsave("figures/compare_spike_ins_atcc/target_offtarget_dedup_2025-01-01.png", width = 10, height = 8)

ggsave("figures/manuscript_figures_pdf/Figure_3_2025-01-01.pdf", width = 10, height = 8)

```

### ----nodedup results ----
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=12, fig.width=6}
##extras - no dedup

nodedup_reads <-
  nodedup_viral_reads |>
  filter(Background.sample != "Neg_control") |>
  filter(Background.sample != "Msp_p6") |>
  ggplot(aes(
    x = Viral.load,
    y = log10(count),
    colour = genus
  )) +
  geom_point() +
  geom_smooth(aes(group = genus, linetype = genome_structure), se = FALSE) +
  facet_grid(target ~ Background.sample, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10() +
  ylab("log10(Viral Reads)") +
  scale_color_manual(values = cols)

nodedup_prop_viral <- 
  nodedup_viral_reads |>
  filter(Background.sample != "Neg_control") |>
  filter(Background.sample != "Msp_p6") |>
  ggplot(aes(x = Viral.load, y = prop_viral, colour = genus)) +
  geom_point() +
  geom_smooth(aes(group = genus, linetype = genome_structure), se = FALSE) +
  facet_grid(target ~ Background.sample, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10() +
  scale_y_log10() +
  ylab("Proportion of viral reads") +
  scale_color_manual(values = cols)

nodedup_prop_all <- nodedup_viral_reads |>
  filter(Background.sample != "Neg_control") |>
  filter(Background.sample != "Msp_p6") |>
  ggplot(aes(x = Viral.load, y = prop_total, colour = genus)) +
  geom_point() +
  geom_smooth(aes(group = genus, linetype = genome_structure), se = FALSE) +
  facet_grid(target ~ Background.sample, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10() +
  scale_y_log10() +
  ylab("Proportion of total reads") +
  scale_color_manual(values = cols)

ggarrange(
  nodedup_reads,
  nodedup_prop_all,
  nodedup_prop_viral,
  nrow = 3,
  common.legend = TRUE,
  align = "hv"
)

#ggsave("figures/compare_spike_ins_atcc/target_offtarget_nodedup.pdf",width=8,height=10)

```

### --------------------